---
title: "04_Pooled N model_NEON data"
author: "Christa Torrens"
format: html
    code-fold: true #enables collapsible code blocks
    code-summary: "Show the code"
editor: visual
---

# Loading data and running the stan pooled-light model

This script runs the pooled-light NO3 model for selected NEON aquatic sites. It has different sections for each site, so they can be run individually. DO NOT RENDER this doc - Rmd knitting does not play well w. running STAN models. However, the .qmd navigation pane is very useful for efficiently moving around the code.

First load the required packages:

```{r loading packages, echo=false}

# load packages - EDIT as needed!
library(scales)
library(magrittr) # moved from  'light ea day as vector'
library(tidyverse)
library(lubridate)
library(streamMetabolizer)
library(rstan)
library(tidybayes)
library(GGally)
library(shinystan)
library(zoo)
library(neonUtilities)
library(ggpubr)
library(brms)
library(here) # allows project-based file paths
library(pracma) # for 1 type of light AUC calcs
library(httr) # helps w. the NSRDB light download

options(mc.cores = parallel::detectCores())

```

Then work with the site-level nitrate and light datasets to create the model data list, and run the STAN model. For each site:

-   Load the datasets
-   Assemble the model data
-   Create the data list and call to STAN
-   Save each model fit and data list as Rdata

### BIGC - Upper Big Creek, Fresno, CA

##### Load data files

```{r - load BIGC data files}
#| output: false
#| message: false

###### NB: the local_datetime (and all other posix/datetime columns) reverts to UTC when saved and reloaded; need to re-apply the tz change here. Local_datetime values do not change. 

###### ALSO need to re-create the model_datetime column: on reload, the values DO  change to = UTC-4h instead of local_datetime-4h. 

# making sure the wd is set correctly
setwd(here())

####################### NO3 + real light data ##################

# Need to re-specify the tz for local_datetime *and* re-derive model_datetime: per code in '03_select NEON data...':  the model_datetime column already exists but values changed to the UTC time - 4h... weird but there it is. 

# full dataset
path <- here("N_uptake_NEON/data/neon_data_clean/bigc_clean.csv")
bigc.df <- read_csv(path) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Pacific"), 
         model_datetime = local_datetime - hours(4))  


# hourly dataset
path_h <- here("N_uptake_NEON/data/neon_data_clean/bigc_hourly_clean.csv")
bigc.df.h <- read_csv(path_h) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Pacific"), 
         model_datetime = local_datetime - hours(4))  

```

##### Visualize NO3 data if needed

```{r - bigc - visualize data, plot, clip as needed}
#low <- 175
#high <- 186
#cliplist <- c(135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147)  # 135-147 = May 15-27, 2019
  
bigc.plot <- bigc.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  #filter(model_day != cliplist) %>%
  ggplot(aes(x=local_datetime, y=surfWaterNitrateMean)) + 
  geom_point() + 
  #geom_line() + 
  labs(x="Date", y=expression("NO"[3]~"(umol/L)"), title= "Big Creek nitrate") +
  theme_bw()

quartz()
bigc.plot

```

##### Assemble data for the model

```{r BIGC - assemble data}
#| output: false
#| message: false

###############  Modeled (ideal) light from streamMetabolizer  ##################

############ Check sunrise/sunset times here: https://gml.noaa.gov/grad/solcalc/
#   ~ 5:30 in June - but the satellite light has data at 4a in FEBRUARY

########  BIGC Coordinates
lat <- 37.05767
lon <- -119.25538

########  Calculate solar time and light, add as columns ____________________

# streamMetabolizer modeled sunlight units = photon flux density (µmol photons m⁻² s⁻¹)

bigc.df.h <- bigc.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  mutate(solar_time = streamMetabolizer::calc_solar_time(local.time=local_datetime, longitude=lon),
         mod_light = streamMetabolizer::calc_light(solar.time=solar_time, latitude=lat,longitude=lon), 
         hours = hour(local_datetime))

######## Visualize light data ___________________________________________________

plot(bigc.df.h$model_datetime, bigc.df.h$mod_light, type='l') #daily rises and falls

######## Determine partial model days ___________________________________________

head(bigc.df.h$model_jday) # 2021_32 is the first model day (complete)
tail(bigc.df.h$model_jday) #2022_364 is the last model day (complete)

bigc.df.h$yr_jday
# how many hours are in the last model day?  Just 20. 
bigc.df.h %>%
  filter(model_jday == 32) %>%
  summarize(count = n())

#lastday <- bigc.df.19h$model_day[]  # add index # here if further clipping is needed

######## Create ideal sumlight data from full dataset _____________________

bigc.sumlightIdeal.df.h <- bigc.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs = sum(mod_light != 0), 
              sum_of_light = sum(mod_light),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlight.trapz = trapz(hours, mod_light)) #same as sum_of_light

# no longer need to remove days from either dataframe, since grouping by model day. 
View(bigc.sumlight.df.h)

summary(bigc.sumlight.df.h)
head(bigc.sumlight.df.h)  # 2021-32
tail(bigc.sumlight.df.h)  # 2022-364
# bigc.df.19h <- bigc.df.19h %>%  # also remove that day from the main df 
#   filter(model_day != 226) 
#   

# Check for days w. 0 hours of light (usually at end) and add to 'darkdays'; use this to remove same days from the real light data. 
which(bigc.sumlight.df.h$light.hrs == 0) #none!
bigc.sumlight.df.h$model_day[which(bigc.sumlight.df.h$light.hrs == 0)] #227 - get the model_day for 0 light

# darkdays <- 227 # c(227)
# 
# sumlight.h <- sumlight.h.df[!(sumlight.h.df$light.hrs == 0),]
# 
# bigc.df.19h <- bigc.df.19h[!(bigc.df.19h$model_day %in% darkdays),]

# tail(bigc.df.19h$model_day)

# get the updated range of model days

  
# How does it look?  => work on that...
# plot(bigc.sumlight.df.h$yr_jday, bigc.sumlight.df.h$sumlight.trapz)  ## ~ 14000 - 20000

## Standardize terms for model
light <- bigc.df.h$mod_light    
sumlight.ideal <- bigc.sumlight.df.h$sumlight.trapz


########  Real daily light (from NSRDB/ satellite) _____________________________

## these data are included in the bigc.df and bigc.df.h datasets; clip if/as  needed to match the steps above

# First remove model days to match sumlight.h ####### NOT NEEDED ##########
# bigc19_sat_sumlight_clip <- bigc19_sat_sumlight %>% 
#   filter(model_day >= low_cut & model_day <= high_cut) 
# %>%
#   filter()  # trim further, if needed, to match model dataset. 

# View(bigc19_sat_sumlight_clip) 

bigc.sumlightReal.df.h <- bigc.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs.Real = sum(GHI_wm2 != 0), 
              sum_of_light.Real = sum(GHI_wm2),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlightReal.trapz = trapz(hours, GHI_wm2))


# standardize for model
sumlight.real <-  bigc.sumlightReal.df.h$sumlightReal.trapz


########  Other elements for the data list  ___________________________________

set.seed(0220)

nday <-  198 # 39

N_init <- bigc.df.h$surfWaterNitrateMean[1] # 3.7 umol l-1
concMA <- matrix(unlist(bigc.df.h$surfWaterNitrateMean), ncol = nday, byrow = FALSE)

Ne_meanprior <- mean(bigc.df.h$surfWaterNitrateMean) # 3.027
Ne_sdprior <- sd(bigc.df.h$surfWaterNitrateMean)     # 1.340

z <- 0.2 #depth in m based on Kelly Aho's dataset - constant for now
zMA<- matrix(z, ncol = nday, nrow=24) #depth at each timestep as matrix (col=days, row= hours/day)

lightMA <- matrix(light, nrow=24)




```

##### List data and call to stan

```{r - BIGC list data and call to stan}



# data list for STAN
data <- list(T=24,D=nday,deltat=1/24,lightMA=lightMA,sumlightIdeal=sumlight.ideal,sumlightReal=sumlight.real,zMA=zMA,concMA=concMA, Ne_meanprior=Ne_meanprior, Ne_sdprior=Ne_sdprior)

# Fit the model
fit.bigc19 <-  stan("N_uptake_NEON/pooled-L_all.stan", data = data,  iter = 2000, chains = 4, control = list(max_treedepth = 15))

print(fit.bigc19)

```

##### Save the model fit as Rdata

```{r - save BIGC model fit}
#| output: false
#| message: false

path <- here("N_uptake_NEON/data/model_fits/cari.fit.rds")
saveRDS(fit.cari, path)

path2 <- here("N_uptake_NEON/data/model_datalist/cari.data.rds")
saveRDS(data, path2)

# To reload: 
# fit.cari <- readRDS(path)
data.cari <- readRDS(path2)

```

### CARI - Caribou Creek, Fairbanks North Star, AK

##### Load data files

```{r - load CARI data}
#| output: false
#| message: false

###### NB: the local_datetime (and all other posix/datetime columns) reverts to UTC when saved and reloaded; need to re-apply the tz change here. Local_datetime values do not change. 

###### ALSO need to re-create the model_datetime column: on reload, the values DO  change to = UTC-4h instead of local_datetime-4h. 

# making sure the wd is set correctly
setwd(here())

####################### NO3 + real light data ##################

# Need to re-specify the tz for local_datetime *and* re-derive model_datetime: per code in '03_select NEON data...':  the model_datetime column already exists but values changed to the UTC time - 4h... weird but there it is. 

# full dataset
path <- here("N_uptake_NEON/data/neon_data_clean/cari_clean.csv")
cari.df <- read_csv(path) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


# hourly dataset
path_h <- here("N_uptake_NEON/data/neon_data_clean/cari_hourly_clean.csv")
cari.df.h <- read_csv(path_h) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


```

##### Visualize NO3 data if needed

```{r - CARI - visualize data, plot, clip as needed}
#low <- 175
#high <- 186
#cliplist <- c(135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147)  # 135-147 = May 15-27, 2019
  
cari.plot <- cari.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  #filter(model_day != cliplist) %>%
  ggplot(aes(x=local_datetime, y=surfWaterNitrateMean)) + 
  geom_point() + 
  #geom_line() + 
  labs(x="Date", y=expression("NO"[3]~"(umol/L)"), title= "West St. Louis Creek nitrate") +
  theme_bw()

quartz(width = 10, height = 4.5)
cari.plot

```

##### Assemble data for the model

```{r - CARI - assemble data}

###############  Modeled (ideal) light from streamMetabolizer  ##################

############ Check sunrise/sunset times here: https://gml.noaa.gov/grad/solcalc/
#   ~ 5:30 in June - but the satellite light has data at 4a in FEBRUARY

########  CARI Coordinates
lat <- 39.890673
lon <- -105.911297

########  Calculate solar time and light, add as columns ____________________

# streamMetabolizer modeled sunlight units = photon flux density (µmol photons m⁻² s⁻¹)

cari.df.h <- cari.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  mutate(solar_time = streamMetabolizer::calc_solar_time(local.time=local_datetime, longitude=lon),
         mod_light = streamMetabolizer::calc_light(solar.time=solar_time, latitude=lat,longitude=lon), 
         hours = hour(local_datetime))

######## Visualize light data ___________________________________________________

plot(cari.df.h$model_datetime, cari.df.h$mod_light, type='l') #daily rises and falls

######## Determine partial model days ___________________________________________

head(cari.df.h$model_jday) # 2021_32 is the first model day (complete)
tail(cari.df.h$model_jday) #2022_364 is the last model day (complete)

cari.df.h$yr_jday
# how many hours are in the last model day?  Just 20. 
cari.df.h %>%
  filter(model_jday == 32) %>%
  summarize(count = n())

#lastday <- bigc.df.19h$model_day[]  # add index # here if further clipping is needed

######## Create ideal sumlight data from full dataset _____________________

cari.sumlightIdeal.df.h <- cari.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs = sum(mod_light != 0), 
              sum_of_light = sum(mod_light),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlight.trapz = trapz(hours, mod_light)) #same as sum_of_light

# no longer need to remove days from either dataframe, since grouping by model day. 
View(cari.sumlight.df.h)

summary(cari.sumlight.df.h)
head(cari.sumlight.df.h)  # 2021-32
tail(cari.sumlight.df.h)  # 2022-364
# bigc.df.19h <- bigc.df.19h %>%  # also remove that day from the main df 
#   filter(model_day != 226) 
#   

# Check for days w. 0 hours of light (usually at end) and add to 'darkdays'; use this to remove same days from the real light data. 
which(cari.sumlight.df.h$light.hrs == 0) #none!
cari.sumlight.df.h$model_day[which(cari.sumlight.df.h$light.hrs == 0)] #227 - get the model_day for 0 light

# darkdays <- 227 # c(227)
# 
# sumlight.h <- sumlight.h.df[!(sumlight.h.df$light.hrs == 0),]
# 
# bigc.df.19h <- bigc.df.19h[!(bigc.df.19h$model_day %in% darkdays),]

# tail(bigc.df.19h$model_day)

# get the updated range of model days

  
# How does it look?  => work on that...
# plot(cari.sumlight.df.h$yr_jday, cari.sumlight.df.h$sumlight.trapz)  ## ~ 14000 - 20000

## Standardize terms for model
light <- cari.df.h$mod_light    #bigc.df.19h$light
sumlight.ideal <- cari.sumlight.df.h$sumlight.trapz


########  Real daily light (from NSRDB/ satellite) _____________________________

## these data are included in the cari.df and cari.df.h datasets; clip if/as  needed to match the steps above

# First remove model days to match sumlight.h ####### NOT NEEDED ##########
# bigc19_sat_sumlight_clip <- bigc19_sat_sumlight %>% 
#   filter(model_day >= low_cut & model_day <= high_cut) 
# %>%
#   filter()  # trim further, if needed, to match model dataset. 

# View(bigc19_sat_sumlight_clip) 

cari.sumlightReal.df.h <- cari.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs.Real = sum(GHI_wm2 != 0), 
              sum_of_light.Real = sum(GHI_wm2),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlightReal.trapz = trapz(hours, GHI_wm2))


# standardize for model
sumlight.real <-  cari.sumlightReal.df.h$sumlightReal.trapz


########  Other elements for the data list  ___________________________________

set.seed(0220)

nday <-  198 # 39

N_init <- cari.df.h$surfWaterNitrateMean[1] # 3.7 umol l-1
concMA <- matrix(unlist(cari.df.h$surfWaterNitrateMean), ncol = nday, byrow = FALSE)

Ne_meanprior <- mean(cari.df.h$surfWaterNitrateMean) # 3.027
Ne_sdprior <- sd(cari.df.h$surfWaterNitrateMean)     # 1.340

z <- 0.2 #depth in m based on Kelly Aho's dataset - constant for now
zMA<- matrix(z, ncol = nday, nrow=24) #depth at each timestep as matrix (col=days, row= hours/day)

lightMA <- matrix(light, nrow=24)


```

##### List data and call to stan

```{r - CARI list data and call to stan}

# data list for STAN
data <- list(T=24,D=nday,deltat=1/24,lightMA=lightMA,sumlightIdeal=sumlight.ideal,sumlightReal=sumlight.real,zMA=zMA,concMA=concMA, Ne_meanprior=Ne_meanprior, Ne_sdprior=Ne_sdprior)

# Fit the model
fit.cari <-  stan("N_uptake_NEON/pooled-L_all.stan", data = data,  iter = 2000, chains = 4, control = list(max_treedepth = 15))

summary(fit.cari)
print(fit.cari, pars=c("sigma", "sigma_U", "b0", "b1"))

# save the model fit summary
# p.bigc19 <- summary(fit.bigc19)
# p.bigc19

```

##### Save the model fit and model data as Rdata

```{r - save CARI model fit and data}
#| output: false
#| message: false

path <- here("N_uptake_NEON/data/model_fits/cari.fit.rds")
saveRDS(fit.cari, path)

path2 <- here("N_uptake_NEON/data/model_datalist/cari.data.rds")
saveRDS(data, path2)

# To reload: 
# fit.cari <- readRDS(path)
data.cari <- readRDS(path2)

```

### CUPE - Rio Cupeyes, San German Municipio, PR

##### Load data files

```{r - load CUPE data}
#| output: false
#| message: false

###### NB: the local_datetime (and all other posix/datetime columns) reverts to UTC when saved and reloaded; need to re-apply the tz change here. Local_datetime values do not change. 

###### ALSO need to re-create the model_datetime column: on reload, the values DO  change to = UTC-4h instead of local_datetime-4h. 

# making sure the wd is set correctly
setwd(here())

####################### NO3 + real light data ##################

# Need to re-specify the tz for local_datetime *and* re-derive model_datetime: per code in '03_select NEON data...':  the model_datetime column already exists but values changed to the UTC time - 4h... weird but there it is. 

# full dataset
path <- here("N_uptake_NEON/data/neon_data_clean/cupe_clean.csv")
cupe.df <- read_csv(path) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


# hourly dataset
path_h <- here("N_uptake_NEON/data/neon_data_clean/cupe_hourly_clean.csv")
cupe.df.h <- read_csv(path_h) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


```

##### Visualize NO3 data if needed

```{r - CUPE - visualize data, plot, clip as needed}
#low <- 175
#high <- 186
#cliplist <- c(135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147)  # 135-147 = May 15-27, 2019
  
cupe.plot <- cupe.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  #filter(model_day != cliplist) %>%
  ggplot(aes(x=local_datetime, y=surfWaterNitrateMean)) + 
  geom_point() + 
  #geom_line() + 
  labs(x="Date", y=expression("NO"[3]~"(umol/L)"), title= "West St. Louis Creek nitrate") +
  theme_bw()

quartz(width = 10, height = 4.5)
cupe.plot

```

##### Assemble data for the model

```{r - CUPE - assemble data}

###############  Modeled (ideal) light from streamMetabolizer  ##################

############ Check sunrise/sunset times here: https://gml.noaa.gov/grad/solcalc/
#   ~ 5:30 in June - but the satellite light has data at 4a in FEBRUARY

########  CUPE Coordinates
lat <- 39.890673
lon <- -105.911297

########  Calculate solar time and light, add as columns ____________________

# streamMetabolizer modeled sunlight units = photon flux density (µmol photons m⁻² s⁻¹)

cupe.df.h <- cupe.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  mutate(solar_time = streamMetabolizer::calc_solar_time(local.time=local_datetime, longitude=lon),
         mod_light = streamMetabolizer::calc_light(solar.time=solar_time, latitude=lat,longitude=lon), 
         hours = hour(local_datetime))

######## Visualize light data ___________________________________________________

plot(cupe.df.h$model_datetime, cupe.df.h$mod_light, type='l') #daily rises and falls

######## Determine partial model days ___________________________________________

head(cupe.df.h$model_jday) # 2021_32 is the first model day (complete)
tail(cupe.df.h$model_jday) #2022_364 is the last model day (complete)

cupe.df.h$yr_jday
# how many hours are in the last model day?  Just 20. 
cupe.df.h %>%
  filter(model_jday == 32) %>%
  summarize(count = n())

#lastday <- bigc.df.19h$model_day[]  # add index # here if further clipping is needed

######## Create ideal sumlight data from full dataset _____________________

cupe.sumlightIdeal.df.h <- cupe.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs = sum(mod_light != 0), 
              sum_of_light = sum(mod_light),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlight.trapz = trapz(hours, mod_light)) #same as sum_of_light

# no longer need to remove days from either dataframe, since grouping by model day. 
View(cupe.sumlight.df.h)

summary(cupe.sumlight.df.h)
head(cupe.sumlight.df.h)  # 2021-32
tail(cupe.sumlight.df.h)  # 2022-364
# bigc.df.19h <- bigc.df.19h %>%  # also remove that day from the main df 
#   filter(model_day != 226) 
#   

# Check for days w. 0 hours of light (usually at end) and add to 'darkdays'; use this to remove same days from the real light data. 
which(cupe.sumlight.df.h$light.hrs == 0) #none!
cupe.sumlight.df.h$model_day[which(cupe.sumlight.df.h$light.hrs == 0)] #227 - get the model_day for 0 light

# darkdays <- 227 # c(227)
# 
# sumlight.h <- sumlight.h.df[!(sumlight.h.df$light.hrs == 0),]
# 
# bigc.df.19h <- bigc.df.19h[!(bigc.df.19h$model_day %in% darkdays),]

# tail(bigc.df.19h$model_day)

# get the updated range of model days

  
# How does it look?  => work on that...
# plot(cupe.sumlight.df.h$yr_jday, cupe.sumlight.df.h$sumlight.trapz)  ## ~ 14000 - 20000

## Standardize terms for model
light <- cupe.df.h$mod_light    #bigc.df.19h$light
sumlight.ideal <- cupe.sumlight.df.h$sumlight.trapz


########  Real daily light (from NSRDB/ satellite) _____________________________

## these data are included in the cupe.df and cupe.df.h datasets; clip if/as  needed to match the steps above

# First remove model days to match sumlight.h ####### NOT NEEDED ##########
# bigc19_sat_sumlight_clip <- bigc19_sat_sumlight %>% 
#   filter(model_day >= low_cut & model_day <= high_cut) 
# %>%
#   filter()  # trim further, if needed, to match model dataset. 

# View(bigc19_sat_sumlight_clip) 

cupe.sumlightReal.df.h <- cupe.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs.Real = sum(GHI_wm2 != 0), 
              sum_of_light.Real = sum(GHI_wm2),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlightReal.trapz = trapz(hours, GHI_wm2))


# standardize for model
sumlight.real <-  cupe.sumlightReal.df.h$sumlightReal.trapz


########  Other elements for the data list  ___________________________________

set.seed(0220)

nday <-  198 # 39

N_init <- cupe.df.h$surfWaterNitrateMean[1] # 3.7 umol l-1
concMA <- matrix(unlist(cupe.df.h$surfWaterNitrateMean), ncol = nday, byrow = FALSE)

Ne_meanprior <- mean(cupe.df.h$surfWaterNitrateMean) # 3.027
Ne_sdprior <- sd(cupe.df.h$surfWaterNitrateMean)     # 1.340

z <- 0.2 #depth in m based on Kelly Aho's dataset - constant for now
zMA<- matrix(z, ncol = nday, nrow=24) #depth at each timestep as matrix (col=days, row= hours/day)

lightMA <- matrix(light, nrow=24)


```

##### List data and call to stan

```{r - CUPE list data and call to stan}

# data list for STAN
data <- list(T=24,D=nday,deltat=1/24,lightMA=lightMA,sumlightIdeal=sumlight.ideal,sumlightReal=sumlight.real,zMA=zMA,concMA=concMA, Ne_meanprior=Ne_meanprior, Ne_sdprior=Ne_sdprior)

# Fit the model
fit.cupe <-  stan("N_uptake_NEON/pooled-L_all.stan", data = data,  iter = 2000, chains = 4, control = list(max_treedepth = 15))

summary(fit.cupe)
print(fit.cupe, pars=c("sigma", "sigma_U", "b0", "b1"))

# save the model fit summary
# p.bigc19 <- summary(fit.bigc19)
# p.bigc19

```

##### Save the model fit and model data as Rdata

```{r - save CUPE model fit and data}
#| output: false
#| message: false


path <- here("N_uptake_NEON/data/model_fits/cupe.fit.rds")
saveRDS(fit.cupe, path)

path2 <- here("N_uptake_NEON/data/model_datalist/cupe.data.rds")
saveRDS(data, path2)

# To reload: 
# fit.cupe <- readRDS(path)
data.cupe <- readRDS(path2)

```

### WLOU - West St Louis Creek, Grand, CO

##### Load data files

```{r - load WLOU data}
#| output: false
#| message: false

###### NB: the local_datetime (and all other posix/datetime columns) reverts to UTC when saved and reloaded; need to re-apply the tz change here. Local_datetime values do not change. 

###### ALSO need to re-create the model_datetime column: on reload, the values DO  change to = UTC-4h instead of local_datetime-4h. 

# making sure the wd is set correctly
setwd(here())

####################### NO3 + real light data ##################

# full dataset
path <- here("N_uptake_NEON/data/neon_data_clean/wlou_clean.csv")
wlou.df <- read_csv(path) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


# hourly dataset
path_h <- here("N_uptake_NEON/data/neon_data_clean/wlou_hourly_clean.csv")
wlou.df.h <- read_csv(path_h) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Mountain"), 
         model_datetime = local_datetime - hours(4))  


```

##### Visualize NO3 data if needed

```{r - WLOU - visualize data, plot, clip as needed}
#low <- 175
#high <- 186
#cliplist <- c(135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147)  # 135-147 = May 15-27, 2019
  
wlou.plot <- wlou.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  #filter(model_day != cliplist) %>%
  ggplot(aes(x=local_datetime, y=surfWaterNitrateMean)) + 
  geom_point() + 
  #geom_line() + 
  labs(x="Date", y=expression("NO"[3]~"(umol/L)"), title= "West St. Louis Creek nitrate") +
  theme_bw()

quartz(width = 10, height = 4.5)
wlou.plot

```

##### Assemble data for the model

```{r - WLOU - assemble data}

###############  Modeled (ideal) light from streamMetabolizer  ##################

############ Check sunrise/sunset times here: https://gml.noaa.gov/grad/solcalc/
#   ~ 5:30 in June - but the satellite light has data at 4a in FEBRUARY

########  WLOU Coordinates
lat <- 39.890673
lon <- -105.911297

########  Calculate solar time and light, add as columns ____________________

# streamMetabolizer modeled sunlight units = photon flux density (µmol photons m⁻² s⁻¹)

wlou.df.h <- wlou.df.h %>%
  #filter(model_day >= low & model_day <= high) %>%
  mutate(solar_time = streamMetabolizer::calc_solar_time(local.time=local_datetime, longitude=lon),
         mod_light = streamMetabolizer::calc_light(solar.time=solar_time, latitude=lat,longitude=lon), 
         hours = hour(local_datetime))

######## Visualize light data ___________________________________________________

plot(wlou.df.h$model_datetime, wlou.df.h$mod_light, type='l') #daily rises and falls

######## Determine partial model days ___________________________________________

head(wlou.df.h$model_jday) # 2021_32 is the first model day (complete)
tail(wlou.df.h$model_jday) #2022_364 is the last model day (complete)

wlou.df.h$yr_jday
# how many hours are in the last model day?  Just 20. 
wlou.df.h %>%
  filter(model_jday == 32) %>%
  summarize(count = n())

#lastday <- bigc.df.19h$model_day[]  # add index # here if further clipping is needed

######## Create ideal sumlight data from full dataset _____________________

wlou.sumlightIdeal.df.h <- wlou.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs = sum(mod_light != 0), 
              sum_of_light = sum(mod_light),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlight.trapz = trapz(hours, mod_light)) #same as sum_of_light

# no longer need to remove days from either dataframe, since grouping by model day. 
View(wlou.sumlight.df.h)

summary(wlou.sumlight.df.h)
head(wlou.sumlight.df.h)  # 2021-32
tail(wlou.sumlight.df.h)  # 2022-364
# bigc.df.19h <- bigc.df.19h %>%  # also remove that day from the main df 
#   filter(model_day != 226) 
#   

# Check for days w. 0 hours of light (usually at end) and add to 'darkdays'; use this to remove same days from the real light data. 
which(wlou.sumlight.df.h$light.hrs == 0) #none!
wlou.sumlight.df.h$model_day[which(wlou.sumlight.df.h$light.hrs == 0)] #227 - get the model_day for 0 light

# darkdays <- 227 # c(227)
# 
# sumlight.h <- sumlight.h.df[!(sumlight.h.df$light.hrs == 0),]
# 
# bigc.df.19h <- bigc.df.19h[!(bigc.df.19h$model_day %in% darkdays),]

# tail(bigc.df.19h$model_day)

# get the updated range of model days

  
# How does it look?  => work on that...
# plot(wlou.sumlight.df.h$yr_jday, wlou.sumlight.df.h$sumlight.trapz)  ## ~ 14000 - 20000

## Standardize terms for model
light <- wlou.df.h$mod_light    #bigc.df.19h$light
sumlight.ideal <- wlou.sumlight.df.h$sumlight.trapz


########  Real daily light (from NSRDB/ satellite) _____________________________

## these data are included in the wlou.df and wlou.df.h datasets; clip if/as  needed to match the steps above

# First remove model days to match sumlight.h ####### NOT NEEDED ##########
# bigc19_sat_sumlight_clip <- bigc19_sat_sumlight %>% 
#   filter(model_day >= low_cut & model_day <= high_cut) 
# %>%
#   filter()  # trim further, if needed, to match model dataset. 

# View(bigc19_sat_sumlight_clip) 

wlou.sumlightReal.df.h <- wlou.df.h %>%
  group_by(year(model_datetime), model_jday) %>%  # with multiple years, need to group by the year-jday column
    summarize(light.hrs.Real = sum(GHI_wm2 != 0), 
              sum_of_light.Real = sum(GHI_wm2),
              # sumlight = sum(mod_light)/(light.hrs),
              # sumlight.h = sum(mod_light)/24, 
              sumlightReal.trapz = trapz(hours, GHI_wm2))


# standardize for model
sumlight.real <-  wlou.sumlightReal.df.h$sumlightReal.trapz


########  Other elements for the data list  ___________________________________

set.seed(0220)

nday <-  198 # 39

N_init <- wlou.df.h$surfWaterNitrateMean[1] # 3.7 umol l-1
concMA <- matrix(unlist(wlou.df.h$surfWaterNitrateMean), ncol = nday, byrow = FALSE)

Ne_meanprior <- mean(wlou.df.h$surfWaterNitrateMean) # 3.027
Ne_sdprior <- sd(wlou.df.h$surfWaterNitrateMean)     # 1.340

z <- 0.2 #depth in m based on Kelly Aho's dataset - constant for now
zMA<- matrix(z, ncol = nday, nrow=24) #depth at each timestep as matrix (col=days, row= hours/day)

lightMA <- matrix(light, nrow=24)


```

##### List data and call to stan

```{r - WLOU list data and call to stan}

# data list for STAN
data <- list(T=24,D=nday,deltat=1/24,lightMA=lightMA,sumlightIdeal=sumlight.ideal,sumlightReal=sumlight.real,zMA=zMA,concMA=concMA, Ne_meanprior=Ne_meanprior, Ne_sdprior=Ne_sdprior)

# Fit the model
fit.wlou <-  stan("N_uptake_NEON/pooled-L_all.stan", data = data,  iter = 2000, chains = 4, control = list(max_treedepth = 15))

# check the summary sigma and beta parameters
summary(fit.wlou)
print(fit.wlou, pars=c("sigma", "sigma_U", "b0", "b1"))

```

##### Save the model fit and model data as Rdata

```{r - save WLOU model fit and data}
#| output: false
#| message: false

path <- here("N_uptake_NEON/data/model_fits/wlou.fit.rds")
saveRDS(fit.wlou, path)

path2 <- here("N_uptake_NEON/data/model_datalist/wlou.data.rds")
saveRDS(data, path2)

# To reload: 
# fit.wlou <- readRDS(path)
# data.wlou <- readRDS(path2)

```

### OLD CODE CHUNKS - glean then delete

```{r old code}

###### NB: the local_datetime (and all other posix/datetime columns) reverts to UTC when saved and reloaded; need to re-apply the tz change here. Local_datetime values do not change. 

###### ALSO need to re-create the model_datetime column: on reload, the values DO  change to = UTC-4h instead of local_datetime-4h. 


# making sure the wd is set correctly
setwd(here())

####################### NO3 data ##################
#~/Documents/R_working/Modelscape/space-time-rivers/N_uptake_NEON/data/neon_data_clean
##### BIGC
bigc19_filepath <- here("N_uptake_NEON/data/neon_data_clean/bigc_2019_74_226.csv")
bigc.df.19h <- read_csv(bigc19_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Pacific"), 
         model_datetime = local_datetime - hours(4))  

bigc21_filepath <- here("N_uptake_NEON/data/neon_data_clean/bigc_2021_60_281.csv")
bigc.df.21h <- read_csv(bigc21_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Pacific"),
         model_datetime = local_datetime - hours(4))

#### CARI

cari19_filepath <- here("N_uptake_NEON/data/neon_data_clean/cari_2019_127_267.csv")
cari.df.19h <- read_csv(cari19_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Alaska"),
         model_datetime = local_datetime - hours(4))

cari20_filepath <- here("N_uptake_NEON/data/neon_data_clean/cari_2020_170_252.csv")
cari.df.20h <- read_csv(cari20_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Alaska"),
         model_datetime = local_datetime - hours(4))

#### KING

king19_filepath <- here("N_uptake_NEON/data/neon_data_clean/king_2019_232_365.csv")
king.df.19h <- read_csv(king19_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Central"),
         model_datetime = local_datetime - hours(4))


king20_filepath <- here("N_uptake_NEON/data/neon_data_clean/king_2020_1_289.csv")
king.df.20h <- read_csv(king20_filepath) %>%
  mutate(local_datetime = with_tz(local_datetime, tzone="US/Central"),
         model_datetime = local_datetime - hours(4))


#### WALK

# walk19_filepath <- here("N_uptake_NEON/data/neon_data_clean/walk_2019_XXXX.csv")
# walk.df.19h <- read_csv(walk_filepath) %>%
# mutate(local_datetime = with_tz(local_datetime, tzone="US/Eastern"),
# model_datetime = local_datetime - hours(4))
# 
# walk20_filepath <- here("N_uptake_NEON/data/neon_data_clean/walk_2020_XXXX.csv")
# walk.df.20h <- read_csv(walk_filepath) %>%
# mutate(local_datetime = with_tz(local_datetime, tzone="US/Eastern"),
# model_datetime = local_datetime - hours(4))


old_obj <- c(bigc19_filepath, bigc21_filepath, cari19_filepath, cari20_filepath, king19_filepath, king20_filepath)
rm(old_obj)

####################### satellite light data ##################

bigc19_sat_sumlight <- read_csv(here("N_uptake_NEON/data/NSRDB_data_clean/bigc2019_realsumlight_73_227.csv"))

bigc21_sat_sumlight <- read_csv(here("N_uptake_NEON/data/NSRDB_data_clean/bigc2021_realsumlight_60_281.csv"))

#cari19_sat_sumlight <- read_csv()

#cari20_sat_sumlight <- read_csv()

king19_sat_sumlight <- read_csv(here("N_uptake_NEON/data/NSRDB_data_clean/king2019_realsumlight_232_365.csv"))

king20_sat_sumlight <- read_csv(here("N_uptake_NEON/data/NSRDB_data_clean/king2020_realsumlight_1_289.csv"))

# walk19_sat_sumlight <- read_csv()
# 
# walk22_sat_sumlight <- read_csv()


```
